### 2022-02-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@59ea7282a67f66615b11db9c9a0529c061308c24 🚀

Ecological networks are a backbone for key ecological and evolutionary
processes; yet enumerating all of the interactions they contain is a
daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
species richness TODO. Recent contributions to the field of ecological
network prediction (Strydom et al., 2021; \textbf{Becker2021OptPre?})
highlight that although interactions can be predicted by adding
ecologically relevant information (in the form of, \emph{e.g.} traits),
we do not have robust guidelines as to how the predictive ability of
these models should be evaluated, nor about how the models should be
trained. Here, by relying on simple derivations and a series of
simulations, we formulate a number of such guidelines, specifically for
the case of binary classifiers derived from thresholded values.
Specifically, we conduct an investigation of the models in terms of
their skill (ability to make the right prediction), bias (trends towards
systematically over-predicting one class), class imbalance (the relative
number of cases representing interactions), and show how these effects
interact. We conclude on the fact that models with the best
interaction-scale predictive score do not necessarily result in the most
accurate representation of the network.

@@

There is an immense diversity of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013). The
choice of metrics with relevance to class-imbalanced problems is
fundamental, because as Japkowicz (2013) unambiguously concluded,
``relatively robust procedures used for unskewed data can break down
miserably when the data is skewed.'' Following Japkowicz (2013), we
focus on two ranking metrics (the areas under the Receiving Operator
Characteristic and Precision Recall curves), and three threshold metrics
(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
show early on that it has undesirable properties).

@@

The dataset used for numerical experiments is composed of 64000 such
\((\xi, \nu)\) pairs, on which four machines are trained: a decision
tree regressor, a boosted regression tree, a ridge regressor, and a
random forest regressor. All models were taken from the \texttt{MLJ.jl}
package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
(Bezanson et al., 2017). All machines use the default parameterization;
this is an obvious deviation from best practices, as the hyperparameters
of any machine require training before its application on a real
dataset. As we use 64000 such datasets, this would require 256000 unique
instances of tweaking the hyperparameters, which is not realistic.
Therefore, we assume that the default parameterizations are comparable
across networks. All machines return a quantitative prediction, usually
(but not necessarily) in \([0,1]\), which is proportional (but not
necessarily linearly) to the probability of an interaction between \(i\)
and \(j\).

In order to pick the best adjacency matrix for a given trained machine,
we performed a thresholding approach using 500 steps on predictions from
the testing set, and picking the threshold that maximized Youden's
informedness, which is usually the optimized target for imbalanced
classification. During the thresholding step, we measured the area under
the receiving-operator characteristic (ROC-AUC) and precision-recall
(PR-AUC) curves, as measures of overall performance over the range of
returned values. We report the ROC-AUC and PR-AUC, as well as a suite of
other measures as introduced in the next section, for the best
threshold. The ensemble model was generated by summing the predictions
of all component models on the testing set (ranged in \([0,1]\)), then
put through the same thresholding process. The complete code to run the
simulations is given as an appendix; running the final simulation
required 4.8 core days (approx. 117 hours).

After the simulations were completed, we removed all runs (\emph{i.e.}
pairs of \(\xi\) and \(\nu\)) for which at least one of the following
conditions was met: the accuracy was 0, the true positive or true
negative rates were 0, the connectance was larger than 0.25. This
removes both the obviously failed model runs, and the networks that are
more densely connected compared to the connectance of empirical food
webs (and are therefore less difficult to predict, being less
imbalanced; preliminary analyses of data with a connectance larger than
3 revealed that all machines reached consistently high performance).

@@

>>>>

Ecological networks are a backbone for key ecological and evolutionary
processes; yet enumerating all of the interactions they contain is a
daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
species richness TODO. Recent contributions to the field of ecological
network prediction (Pichler et al., 2020; Strydom et al., 2021;
\textbf{Becker2021OptPre?}) highlight that although interactions can be
predicted by adding ecologically relevant information (in the form of,
\emph{e.g.} traits), we do not have robust guidelines as to how the
predictive ability of these models should be evaluated, nor about how
the models should be trained. Here, by relying on simple derivations and
a series of simulations, we formulate a number of such guidelines,
specifically for the case of binary classifiers derived from thresholded
values. Specifically, we conduct an investigation of the models in terms
of their skill (ability to make the right prediction), bias (trends
towards systematically over-predicting one class), class imbalance (the
relative number of cases representing interactions), and show how these
effects interact. We conclude on the fact that models with the best
interaction-scale predictive score do not necessarily result in the most
accurate representation of the network.

@@

There is an immense diversity of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013). The
choice of metrics with relevance to class-imbalanced problems is
fundamental, because as Japkowicz (2013) unambiguously concluded,
``relatively robust procedures used for unskewed data can break down
miserably when the data is skewed.'' Following Japkowicz (2013), we
focus on two ranking metrics (the areas under the Receiver Operating
Characteristic and Precision Recall curves), and three threshold metrics
(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
show early on that it has undesirable properties).

@@

The dataset used for numerical experiments is composed of 64000 such
\((\xi, \nu)\) pairs, on which four machines are trained: a decision
tree regressor, a boosted regression tree, a ridge regressor, and a
random forest regressor. All models were taken from the \texttt{MLJ.jl}
package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
(Bezanson et al., 2017). All machines use the default parameterization;
this is an obvious deviation from best practices, as the hyperparameters
of any machine require training before its application on a real
dataset. As we use 64000 such datasets, this would require 256000 unique
instances of tweaking the hyperparameters, which is not realistic.
Therefore, we assume that the default parameterizations are comparable
across networks. All machines return a quantitative prediction, usually
(but not necessarily) in \([0,1]\), which is proportional (but not
necessarily linearly) to the probability of an interaction between \(i\)
and \(j\). Nevertheless, the ROC-AUC and PR-AUC (and therefore the
thresholds) can be measured by integrating over the domain of the values
return by each machine.

In order to pick the best adjacency matrix for a given trained machine,
we performed a thresholding approach using 500 steps on predictions from
the testing set, and picking the threshold that maximized Youden's
informedness. During the thresholding step, we measured the area under
the receiver operating characteristic (ROC-AUC) and precision-recall
(PR-AUC) curves, as measures of overall performance over the range of
returned values. We report the ROC-AUC and PR-AUC, as well as a suite of
other measures as introduced in the next section, for the best
threshold. The ensemble model was generated by summing the predictions
of all component models on the testing set (ranged in \([0,1]\)), then
put through the same thresholding process. The complete code to run the
simulations is available at \texttt{10.17605/OSF.IO/JKEWD}; the
simulations require approx. 5 core days.

After the simulations were completed, we removed all runs (\emph{i.e.}
pairs of \(\xi\) and \(\nu\)) for which at least one of the following
conditions was met: the accuracy was 0, the true positive or true
negative rates were 0, the connectance was larger than 0.25. This
removes both the obviously failed model runs, and the networks that are
more densely connected compared to the connectance of empirical food
webs (and are therefore less difficult to predict, being less
imbalanced; preliminary analyses of data with a connectance larger than
0.3 revealed that all machines reached consistently high performance).

@@

\leavevmode\hypertarget{ref-Pichler2020MacLea}{}%
Pichler, M., Boreux, V., Klein, A.-M., Schleuning, M., \& Hartig, F.
(2020). Machine learning algorithms to infer trait-matching and predict
species interactions in ecological networks. \emph{Methods in Ecology
and Evolution}, \emph{11}(2), 281--293.
\url{https://doi.org/10.1111/2041-210X.13329}
